Approximation as Differentiation

### 2025-06-09 main/first-order.tex
Some minor improvements to families construction.

\noindent The $\partial X$ notation used here is intended to suggest the connection between the Grothendieck
construction and the idea of tangent spaces and derivatives. For any $(f, \partial f): (X, \partial X) \to (Y,
\partial Y)$ and $(g, \partial g): (Y, \partial Y) \to (Z, \partial Z)$ in $\Grothendieck{}T$, the composition
$(g, \partial g) \comp (f, \partial f): (X, \partial X) \to (Z, \partial Z)$ is given by $(g \comp f,
T(f)(\partial g) \comp \partial f)$, with the functor $T(f)$ acting to ``reindex'' $\partial g$ along $f$,
yielding a morphism $T(f)(\partial g): T(f)(\partial Y) \to T(f)(T(g)(\partial Z)) = T(g \comp f)(\partial
Z)$.

@@

For CHAD we are interested in the specific case of the Grothendieck construction where $\cat{C} = \Set$. In
that case our indexed categories $T$ have the form $\Func{-}{\cat{C}}: \Set^\op \to \Cat$, associating to each
set $X$ the functor category $\Func{X}{\cat{C}}$ which has as objects all functors $F: X \to \cat{C}$, namely
$X$-indexed families of objects of $\cat{C}$, and as morphisms $F \to G$, all families of morphisms $\eta(x):
F(x) \to G(x)$ in $\cat{C}$, with naturality trivial because $X$ has only identity morphisms.

Like any functor category $\Func{X}{\cat{C}}$ inherits limits and colimits pointwise from its codomain.

\begin{proposition}
If $\cat{C}$ has limits (resp.~colimits) then $\Func{X}{\cat{C}}$ has limits (resp.~colimits) computed
pointwise.
\end{proposition}

\begin{definition}[Reindexing]
For any function $f: X \to Y$ define the \emph{reindexing} functor $\reindex{-}{f}: \Func{Y}{\cat{C}} \to
\Func{X}{\cat{C}}$ which sends any $Y$-indexed family $F$ over $\cat{C}$ to the $X$-indexed family
$\reindex{F}{f} = F \comp f$ (regarding $f$ as a functor between $X$ and $Y$ as discrete categories) and any
family of morphisms $\eta: F \to G$ to $\reindex{\eta}{f}: \reindex{F}{f} \to \reindex{G}{f}$ with
$\reindex{\eta}{f}(x) = \eta(f(x))$. \todo{Should be clearer that $\reindex{-}{f}$ is actually
$\Func{f}{\cat{C}}$.}
\end{definition}

>>>>

\noindent The $\partial X$ notation is intended to suggest the connection between the Grothendieck
construction and the idea of tangent spaces and derivatives. For any $(f, \partial f): (X, \partial X) \to (Y,
\partial Y)$ and $(g, \partial g): (Y, \partial Y) \to (Z, \partial Z)$ in $\Grothendieck{}T$, the composition
$(g, \partial g) \comp (f, \partial f): (X, \partial X) \to (Z, \partial Z)$ is given by $(g \comp f,
T(f)(\partial g) \comp \partial f)$, with the functor $T(f)$ acting to ``reindex'' $\partial g$ along $f$,
yielding a morphism $T(f)(\partial g): T(f)(\partial Y) \to T(f)(T(g)(\partial Z)) = T(g \comp f)(\partial
Z)$.

@@

For CHAD we are interested in the specific case of the Grothendieck construction where $\cat{C} = \Set$ and
each category $T(X) = \Func{X}{\cat{C}}$, the functor category which has as objects all $X$-indexed families
of objects of $\cat{C}$ (regarded as functors $F: X \to \cat{C}$) , and as morphisms $F \to G$, all families
of morphisms $\eta(x): F(x) \to G(x)$ in $\cat{C}$ (with naturality trivial because $X$ has only identity
morphisms).

% Like any functor category $\Func{X}{\cat{C}}$ inherits limits and colimits pointwise from its codomain.
%
% \begin{proposition}
% If $\cat{C}$ has limits (resp.~colimits) then $\Func{X}{\cat{C}}$ has limits (resp.~colimits) computed
% pointwise.
% \end{proposition}

\begin{definition}[Reindexing]
For any function $f: X \to Y$ define the \emph{reindexing} functor $\reindex{-}{f} = \Func{f}{\cat{C}}:
\Func{Y}{\cat{C}} \to \Func{X}{\cat{C}}$ which sends any $Y$-indexed family $F$ over $\cat{C}$ to the
$X$-indexed family $\reindex{F}{f} = F \comp f$ (regarding $f$ as a functor between $X$ and $Y$ as discrete
categories) and any family of morphisms $\eta: F \to G$ to $\reindex{\eta}{f}: \reindex{F}{f} \to
\reindex{G}{f}$ with $\reindex{\eta}{f}(x) = \eta(f(x))$.
\end{definition}
